---
title: 'NNNS Project: Beta and Cox Regression'
author: "Sima and Brody"
date: "2023-12-06"
output: html_document
---


# Load packages

```{r,inclue=FALSE}
library(knitr)
library(kableExtra)
library(mice)
library(gamlss)
library(survival)
library(tidyverse)
library(survminer)
```

# Objective 1:
Are lower pre- or post-op attention scores associated with lower % oral feeds at discharge?

we used all covariates except for other NNNS scores.

Changing 1s to less than 1:

```{r,include=FALSE}
nnns_imputed<- readRDS("C:/Users/u0753531/OneDrive - University of Utah/Desktop/StatPractice/survival_project/scripts/nnns_imputed.rds")
data=complete(nnns_imputed, 1)
names(data)
attach(data)
response=Percent_of_feeds_taken_by_mouth_at_discharge

# Check if there are any non-zero values in the response
non_zero_indices <- which(response != 0)

if (length(non_zero_indices) > 0) {
  # Apply the transformation to non-zero values
  response[non_zero_indices] <- ((response[non_zero_indices]*(length(non_zero_indices)-1)) + 0.5) / length(non_zero_indices)
}


```

```{r,include=FALSE}

# Define a mapping of old variable names to new, shorter names
variable_mapping <- c(
  "Female" = "Female",
  "Age_at_Surgery_days" = "Age",
  "Cardiac_Anatomy_collapsed" = "Cardiac",
  "Genetic_Syndrome_or_Chromosomal_Abnormality" = "Genetic",
  "Pre_Op_NNNS_attention_score" = "Attention_score",
  "Pre_Op_NNNS_handling_score" = "Handling_score",
  "Pre_Op_NNNS_Quality_of_Movement_Score" = "Movement_score",
  "Pre_Op_NNNS_Regulation_Score" = "Regulation_score",
  "Pre_Op_NNNS_Non_Optimal_Reflexes_Score" = "Reflexes_score",
  "Pre_Op_NNNS_Stress_Score" = "Stress_score",
  "Pre_Op_NNNS_Arousal_Score" = "Arousal_score",
  "Pre_Op_NNNS_Hypertonic_Score" = "Hypertonic_score",
  "Pre_Op_NNNS_Hypotonic_Score" = "Hypotonic_score",
  "Pre_Op_NNNS_Asymmetry_Score" = "Asymmetry_score",
  "Pre_Op_NNNS_Excitability_Score" = "Excitability_score",
  "Pre_Op_NNNS_Lethargy_Score" = "Lethargy_score",
  "Cardiac_Anatomy_collapsed"="Cardiac"
)

# Rename the columns of the data frame using the mapping
colnames(data) <- sapply(colnames(data), function(x) ifelse(x %in% names(variable_mapping), variable_mapping[x], x))
attach(data)
Cardiac=as.factor(Cardiac)


```

```{r,include=FALSE}
# Define a mapping of old variable names to new shorter names
variable_mapping <- c(
  "Post_Op_NNNS_attention_score" = "Po_Attention_score",
  "Post_Op_NNNS_handling_score" = "Po_Handling_score",
  "Post_Op_NNNS_Quality_of_Movement_Score" = "Po_Movement_score",
  "Post_Op_NNNS_Regulation_Score" = "Po_Regulation_score",
  "Post_Op_NNNS_Non_Optimal_Reflexes_Score" = "Po_Reflexes_score",
  "Post_Op_NNNS_Stress_Score" = "Po_Stress_score",
  "Post_Op_NNNS_Arousal_Score" = "Po_Arousal_score",
  "Post_Op_NNNS_Hypertonic_Score" = "Po_Hypertonic_score",
  "Post_Op_NNNS_Hypotonic_Score" = "Po_Hypotonic_score",
  "Post_Op_NNNS_Asymmetry_Score" = "Po_Asymmetry_score",
  "Post_Op_NNNS_Excitability_Score" = "Po_Excitability_score",
  "Post_Op_NNNS_Lethargy_Score" = "Po_Lethargy_score"
)
colnames(data) <- sapply(colnames(data), function(x) ifelse(x %in% names(variable_mapping), variable_mapping[x], x))
attach(data)
```


```{r}
# Assuming 'response' is your data for the histogram
hist(Percent_of_feeds_taken_by_mouth_at_discharge, main = 'Histogram of Percentage of Oral Feeds Taken by Mouth at Discharge', xlab = 'Percentage of Oral Feeds', ylab = 'Frequency', col = 'lightblue', border = 'black', breaks = 10)


```
After changing non-zero values to be less than 1:
```{r}
hist(response, main = 'Histogram of Percentage of Oral Feeds Taken by Mouth at Discharge', xlab = 'Percentage of Oral Feeds', ylab = 'Frequency', col = 'lightblue', border = 'black', breaks = 10)
mtext(text = 'Note: Values changed by y(n-1)+0.5/n to avoid having 1', side = 1, line = 4, adj = 0.5, cex = 0.8)

```

```{r}

plot(data$Attention_score, data$response, 
     main = 'Scatter Plot of pre_attention Score vs Response',
     xlab = 'Total Score', ylab = 'Response')



```

```{r}
plot(data$Po_Attention_score, data$response, 
     main = 'Scatter Plot of post_attention Score vs Response',
     xlab = 'Post Attention Score', ylab = 'Response')
```


Analysis of model for pre and post:

First considering pre and post attention scores in one model:

```{r}
complete_model <- gamlss(response ~   Age+Cardiac+Female+ GI_Complication+Premature+Length_of_intubation_days+Length_of_Stay_days+                                    Genetic+Extubation_failure+Attention_score+Po_Attention_score,
                            
    sigma.formula =~1, nu.formula = ~Age+Cardiac+Female+ GI_Complication+Premature+Length_of_intubation_days+Length_of_Stay_days+                                    Genetic+Extubation_failure+Attention_score+Po_Attention_score, family = BEZI, data
  )
summary(complete_model)
```
```{r,include=FALSE}
# Extract mu and nu coefficients and confidence intervals
mu_info <- broom::tidy(exp(coef(complete_model, "mu")))
nu_info <- broom::tidy(exp(coef(complete_model, "nu")))

# Add a column to specify the component
mu_info$Component <- "mu"
nu_info$Component <- "nu"

# Combine mu and nu information
result_table <- rbind(mu_info, nu_info)
a=(exp(confint(complete_model)))[-14,]


result_table2=data.frame(result_table,a)
library(dplyr)

mu_table <- result_table2 %>%
  filter(Component == "mu") %>%
  as.data.frame()

# Extract nu values
nu_table <- result_table2 %>%
  filter(Component == "nu") %>%
  as.data.frame()

mu_table1=mu_table[,-3]
mu_table2=round(mu_table1[,c(2,3,4)],3)
p_values_mu <- round(c(
0.1719,0.1330,0.7841,0.2907,0.5487,0.8351,0.2805,0.0284 
 ,0.1551,0.0589,0.1649,0.2739,0.6104)
,3)
names(data)
mu_table2=cbind(mu_table2,p_values_mu)

colnames(mu_table2)=c( "Estimated Values", "lower CI","Upper CI","p-value")
rownames(mu_table2)=c("intercept","Age","Single ventricle w/ arch obstruction","CardiacTwo ventricle w/ arch obstruction","Female",
"GI_Complication",
"Premature",
"Length_of_intubation_days",
"Length_of_Stay_days",
"Genetic","Extubation_failure",
"Pre_Attention score","Post_Attention score")

```

table for Odds of >0% oral feeds:
```{r}
html_table <- mu_table2 %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Odds of oral feed when oral feed is> 0%" = 4))

html_table %>%
  row_spec(c(8), extra_css = "background-color: #f2f2f2; font-weight: bold;") %>%
  column_spec(1, extra_css = "font-weight: bold; color: Black;")


```



table for Odds of 0% oral feeds for pre and post scores:
```{r}
nu_table <- result_table2 %>%
  filter(Component == "nu") %>%
  as.data.frame()

nu_table1=nu_table[,-3]
nu_table2=round(nu_table1[,c(2,3,4)],3)

p_values_nu3=round(c(0.198017,0.826825,0.000453 
 ,0.022337 

 ,0.203229,0.094465,0.963463,0.001123 

 ,0.110566,0.430823,0.031817 

 ,0.102482,0.446472),3)
nu_table2=cbind(nu_table2,p_values_nu3)

colnames(nu_table2)=c( "Esrimated Values", "lower CI","Upper CI","P_values")
rownames(nu_table2)=c("intercept","Age","Single ventricle w/ arch obstruction","CardiacTwo ventricle w/ arch obstruction","Female",
"GI_Complication",
"Premature",
"Length_of_intubation_days",
"Length_of_Stay_days",
"Genetic","Extubation_failure","Pre_Attention_score",
"Post_Attention_score")


```

```{r}
html_table <- nu_table2 %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Odds of oral feed when oral feed is 0%" = 4))

html_table %>%
  row_spec(c(3,4,8,11), extra_css = "background-color: #f2f2f2; font-weight: bold;") %>%
  column_spec(1, extra_css = "font-weight: bold; color: Black;")

```


Having separate model for pre and post attention scores:

Pre:
```{r}
complete_model <- gamlss(response ~   Age+Cardiac+Female+ GI_Complication+Premature+Length_of_intubation_days+Length_of_Stay_days+                                    Genetic+Extubation_failure+Attention_score,
                            
    sigma.formula =~1, nu.formula = ~Age+Cardiac+Female+ GI_Complication+Premature+Length_of_intubation_days+Length_of_Stay_days+                                    Genetic+Extubation_failure+Attention_score, family = BEZI, data
  )
summary(complete_model)

```


find the estimated values and confidence interval after exponentiation:


```{r,include=FALSE}
# Extract mu and nu coefficients and confidence intervals
mu_info <- broom::tidy(exp(coef(complete_model, "mu")))
nu_info <- broom::tidy(exp(coef(complete_model, "nu")))

# Add a column to specify the component
mu_info$Component <- "mu"
nu_info$Component <- "nu"

# Combine mu and nu information
result_table <- rbind(mu_info, nu_info)
a=(exp(confint(complete_model)))[-12,]
dim(a)
round(a,2)

result_table2=data.frame(result_table,a)
library(dplyr)

mu_table <- result_table2 %>%
  filter(Component == "mu") %>%
  as.data.frame()

# Extract nu values
nu_table <- result_table2 %>%
  filter(Component == "nu") %>%
  as.data.frame()

mu_table1=mu_table[,-3]
mu_table2=round(mu_table1[,c(2,3,4)],3)
p_values_mu <- round(c(
 0.1484,0.1448,0.9024,0.2495,0.6057,0.9270,0.2918,0.0229 
 ,0.1724,0.0517,0.1688,0.1881)
,3)
names(data)
mu_table2=cbind(mu_table2,p_values_mu)

colnames(mu_table2)=c( "Estimated Values", "lower CI","Upper CI","p-value")
rownames(mu_table2)=c("intercept","Age","Single ventricle w/ arch obstruction","CardiacTwo ventricle w/ arch obstruction","Female",
"GI_Complication",
"Premature",
"Length_of_intubation_days",
"Length_of_Stay_days",
"Genetic","Extubation_failure",
"Pre_Attention_score")

```

table for Odds of >0% oral feeds:
```{r}

html_table <- mu_table2 %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Odds of oral feed when oral feed is> 0%" = 4))

html_table %>%
  row_spec(c(8), extra_css = "background-color: #f2f2f2; font-weight: bold;") %>%
  column_spec(1, extra_css = "font-weight: bold; color: Black;")


```


table for Odds of 0% oral feeds:
```{r}
nu_table <- result_table2 %>%
  filter(Component == "nu") %>%
  as.data.frame()

nu_table1=nu_table[,-3]
nu_table2=round(nu_table1[,c(2,3,4)],3)
p_values_nu <-round( c(
 0.279734,0.793783,0.000499 

 ,0.023769 

 ,0.177770,0.113693,0.992389,0.001088 

 ,0.080149,0.500097,0.032284 
 ,0.117946),3   # Lethargy_score
)
nu_table2=cbind(nu_table2,p_values_nu)


colnames(nu_table2)=c( "Estimated Values", "lower CI","Upper CI","P-values")
rownames(nu_table2)=c("intercept","Age","Single ventricle w/ arch obstruction","CardiacTwo ventricle w/ arch obstruction","Female",
"GI_Complication",
"Premature",
"Length_of_intubation_days",
"Length_of_Stay_days",
"Genetic","Extubation_failure",
"Pre_Attention_score")


```


```{r}
html_table=nu_table2 %>% round(digits = 3) %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Odds of oral feed when oral feed is 0%" = 4))

html_table %>%
  row_spec(c(3,4,8,11), extra_css = "background-color: #f2f2f2; font-weight: bold;") %>%
  column_spec(1, extra_css = "font-weight: bold; color: Black;")

```



Post:
```{r}
complete_model <- gamlss(response~  Age+Cardiac+Female+ GI_Complication+Premature+Length_of_intubation_days+Length_of_Stay_days+                                    Genetic+Extubation_failure+Po_Attention_score,
                            
    sigma.formula =~1, nu.formula = ~Age+Cardiac+Female+ GI_Complication+Premature+Length_of_intubation_days+Length_of_Stay_days+                                    Genetic+Extubation_failure+Po_Attention_score, family = BEZI, data)
summary(complete_model)
```

Exponentiating estimtaes:
```{r}


# Extract mu and nu coefficients and confidence intervals
mu_info <- broom::tidy(exp(coef(complete_model, "mu")))
nu_info <- broom::tidy(exp(coef(complete_model, "nu")))

# Add a column to specify the component
mu_info$Component <- "mu"
nu_info$Component <- "nu"

# Combine mu and nu information
result_table <- rbind(mu_info, nu_info)
a=exp(confint(complete_model))[-13,]
dim(a)

result_table2=data.frame(result_table,a)
library(dplyr)

mu_table <- result_table2 %>%
  filter(Component == "mu") %>%
  as.data.frame()

# Extract nu values
nu_table <- result_table2 %>%
  filter(Component == "nu") %>%
  as.data.frame()

mu_table1=mu_table[,-3]
mu_table2=round(mu_table1[,c(2,3,4)],3)
p_values_mu2 <- round(c(
0.1790,0.0800 
 ,0.5137,0.3750,0.3713,0.8477,0.1435,0.0499 

 ,0.1598,0.0487 
 ,0.1466,0.3691),3   # Po_Lethargy_score
)
mu_table2=cbind(mu_table2,p_values_mu2)





colnames(mu_table2)=c( "Estimated Values", "lower CI","Upper CI","p-values")
rownames(mu_table2)=c("intercept","Age","Single ventricle w/ arch obstruction","CardiacTwo ventricle w/ arch obstruction","Female",
"GI_Complication",
"Premature",
"Length_of_intubation_days",
"Length_of_Stay_days",
"Genetic","Extubation_failure",
"Post_Attention_score")

```
table for Odds of >0% oral feeds for post scores:
```{r}


html_table <- mu_table2 %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Odds of oral feed when oral feed is> 0%" = 4))


html_table %>%
  row_spec(c(10), extra_css = "background-color: #f2f2f2; font-weight: bold;") %>%
  column_spec(1, extra_css = "font-weight: bold; color: Black;")

```

table for Odds of 0% oral feeds for post scores:
```{r}
nu_table <- result_table2 %>%
  filter(Component == "nu") %>%
  as.data.frame()

nu_table1=nu_table[,-3]
nu_table2=round(nu_table1[,c(2,3,4)],3)

p_values_nu2 <-round( c(
 0.171438,0.562887,0.000665 
 ,0.039072 
 ,0.121903,0.080973,0.478396,0.002421 
 ,0.111596,0.428137,0.024077 
 ,0.582392),3   # Po_Lethargy_score
)
nu_table2=cbind(nu_table2,p_values_nu2)

colnames(nu_table2)=c( "Estimated Values", "lower CI","Upper CI","P_values")
rownames(nu_table2)=c("intercept","Age","Single ventricle w/ arch obstruction","CardiacTwo ventricle w/ arch obstruction","Female",
"GI_Complication",
"Premature",
"Length_of_intubation_days",
"Length_of_Stay_days",
"Genetic","Extubation_failure",
"Post_Attention_score")


```

```{r}
html_table <- nu_table2 %>%
  kable("html") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "Odds of oral feed when oral feed is 0%" = 4))

html_table %>%
  row_spec(c(3,4,8,11), extra_css = "background-color: #f2f2f2; font-weight: bold;") %>%
  column_spec(1, extra_css = "font-weight: bold; color: Black;")

```


Summary of Results:
we used three models. One beta regression model included post and pre attention scores. another two models included post and pre attention scores separately. we concluded from all three models that pre and post attention scores are not associated with the lower % oral feeds at discharge.

# Objective 2:

## Cox Regression

```{r}
# Check the structure of your data
str(data)
```

```{r}
# Create a time variable (days to event or censoring) and an event indicator variable (1 for event, 0 for censored)

data <- mutate(data, 
               Time_to_event = ifelse(!is.na(Date_Reaching_Full_PO), 
                                      as.numeric(difftime(as.Date(Date_Reaching_Full_PO), 
                                                          as.Date(Date_PO_feeds_started), 
                                                          units = "days")),
                                      Length_of_Stay_days),
               Event_indicator = ifelse(!is.na(Date_Reaching_Full_PO), 1, 0))

```

### Fit Cox Proportional Hazards models

#### Both Pre- and Postoperative Attention Scores

```{r}
# Univariable

cox_model1 <- coxph(Surv(Time_to_event, Event_indicator) ~ 
                            Attention_score + 
                            Po_Attention_score, 
                          data = data)

summary_cox_model1 <- tidy(cox_model1, conf.int = TRUE, exponentiate = TRUE)

summary(cox_model1)
```

```{r}
# Multivariable
cox_model2 <- coxph(Surv(Time_to_event, Event_indicator) ~ 
                     Age + 
                     Cardiac + 
                     Female + 
                     GI_Complication + 
                     Premature + 
                     Length_of_intubation_days + 
                     Genetic + 
                     Extubation_failure + 
                     Attention_score + 
                     Po_Attention_score,
                   data = data)

summary_cox_model2 <- tidy(cox_model2, conf.int = TRUE, exponentiate = TRUE)

summary(cox_model2)
```
Note: do not think this model should be used since it includes both pre and post attention scores

#### Preoperative Attention Scores

```{r}
# Univariate 
pre_op_uni_model <- coxph(Surv(Time_to_event, Event_indicator) ~ Attention_score, data = data)

summary_pre_op_uni_model <- tidy(pre_op_uni_model, conf.int = TRUE, exponentiate = TRUE)

summary(pre_op_uni_model)
```
Univariable interpretation: 

Preoperative NNNS attention scores were not associated with time to achieve oral feed (HR: 1.158, 95% CI (0.8638, 0.9516), p-value = 0.143). 

```{r}
# Multivariate model with selected co-variates

pre_op_multi_model <- coxph(Surv(Time_to_event, Event_indicator) ~
                              Attention_score + 
                              Age +
                              Cardiac + 
                              Female + 
                              GI_Complication + 
                              Premature + 
                              Length_of_intubation_days + 
                              Genetic +
                              Extubation_failure,
                            data = data)

summary_pre_op_multi_model <- tidy(pre_op_multi_model, conf.int = TRUE, exponentiate = TRUE)

summary(pre_op_multi_model)
```

Multivariable interpretation:  

Attention score appeared to be a marginally good predictor of time to full oral feed; however, this association was not signficant (HR = 1.259,95% CI: ; p-value = 0.086). 

Several other variables showed significant associations: Premature birth (HR = 3.209, p = 0.011); Cardiac conditions (Single ventricle: HR = 0.477, p = 0.019; Two ventricle: HR = 0.520, p = 0.026); Length of intubation days (HR = 0.838, p = 0.007); Extubation failure (hazard ratio = 0.300, p = 0.017)

The overall model demonstrated good predictive accuracy (Concordance = 0.708). Likelihood ratio test (p < 0.001), Wald test (p = 0.0002), and Score (logrank) test (p = 0.0001) indicated that the model is statistically significant.

#### Postoperative Attention Scores
```{r}
# Univariable
post_op_uni_model <- coxph(Surv(Time_to_event, Event_indicator) ~ Po_Attention_score, data = data)

summary_post_op_uni_model <- tidy(post_op_uni_model, conf.int = TRUE, exponentiate = TRUE)

summary(post_op_uni_model)
```

Univariable Analysis:

In the univariable analysis, postoperative NNNS attention scores did not show a statistically significant association with time to full oral feed (HR: 1.148, 95% CI: 0.8713, 0.9044; p = 0.257).

```{r}
# Multivariable

post_op_multi_model <- coxph(Surv(Time_to_event, Event_indicator) ~
                               Po_Attention_score + 
                               Age +
                               Cardiac + 
                               Female + 
                               GI_Complication + 
                               Premature + 
                               Length_of_intubation_days + 
                               Genetic + 
                               Extubation_failure,
                             data = data)

summary_post_op_multi_model <- tidy(post_op_multi_model, conf.int = TRUE, exponentiate = TRUE)

summary(post_op_multi_model)
```

Multivariable Analysis:

In the multivariable analysis, postoperative NNNS attention scores remained non-significant (HR = 0.926; p = 0.574). Several variables showed significant associations:
Premature birth (HR = 4.506, p < 0.001); Cardiac conditions (Single ventricle: HR = 0.412, p = 0.009; Two ventricle: HR = 0.565, p = 0.047); Length of intubation days (HR = 0.872, p = 0.021); Extubation failure (HR = 0.302, p = 0.014)

The overall model exhibited good predictive accuracy (Concordance = 0.707).Likelihood ratio test (p = 3e-05), Wald test (p = 5e-04), and Score (logrank) test (p = 2e-04) indicated that the multivariable model is statistically significant.

#### Graphs

##### Combined

```{r}
# Uni-variable
ggsurvplot(survfit(cox_model1),
           data = data,
           color = "#2E9FDF",
           ggtheme = theme_minimal())
```


```{r}
# Multi-variable
ggsurvplot(survfit(cox_model2),
           data = data,
           color = "#2E9FDF",
           ggtheme = theme_minimal())
```

##### Preoperative Attention Score

```{r}
# Uni-variable
ggsurvplot(survfit(pre_op_uni_model),
           data = data,
           color = "#2E9FDF",
           title = "Pre-Operative Attention Score (Univariable) - Survival Curve ",
           xlab = "Time (in days)",
           ylab = "Failing to Achieve Full Oral Feed Probability",
           risk.table.title = "Number at Risk",
           risk.table.col = "strata",
           risk.table = TRUE,
           ggtheme = theme_minimal())
```


```{r}
# Multi-variable
ggsurvplot(survfit(pre_op_multi_model),
           data = data,
           color = "#2E9FDF",
           ggtheme = theme_minimal())
```

##### Post-operative Attention Score
```{r}
# Uni-variable
ggsurvplot(survfit(post_op_uni_model),
           data = data,
           color = "#2E9FDF",
           title = "Post-Operative Attention Score (Univariable) - Survival Curve ",
           xlab = "Time (in days)",
           ylab = "Failing to Achieve Full Oral Feed Probability",
           risk.table.title = "Number at Risk",
           risk.table.col = "strata",
           risk.table = TRUE,
           ggtheme = theme_minimal())
```


```{r}
# Multi-variable
ggsurvplot(survfit(post_op_multi_model), 
           data = data,
           color = "#2E9FDF",
           ggtheme = theme_minimal())
```